---
title: "WGTS Alignment QC Summary"
author: "CCGCM - Genomics Platform Group"
date: now
date-format: "YYYY-MM-DD HH:mm Z"
execute:
  echo: false
format:
  html:
    toc: false
    toc-expand: 1
    toc-title: Contents
    toc-location: body
    highlight-style: github
    number-sections: false
    link-external-icon: false
    link-external-newwindow: true
    embed-resources: true
    code-copy: true
    code-link: true
    code-fold: true
    code-block-border-left: true
    smooth-scroll: true
    grid:
      body-width: 2800px
params:
  tidy_data: "~/projects/dracarys/inst/reports/wgts-qc/nogit/tidy_data_rds/2025-01-14_wgts.rds"
---

```{r}
#| label: pkgs
#| message: false
#| warning: false

{
  require(dplyr)
  require(dracarys, include.only = "session_info_kable")
  require(DT, include.only = "datatable")
  require(forcats, include.only = "fct_rev")
  require(glue, include.only = "glue")
  require(here, include.only = "here")
  require(plotly, include.only = "ggplotly")
  require(purrr, include.only = "map")
  require(readr, include.only = "read_rds")
  require(stringr, include.only = "str_extract")
  require(tibble, include.only = "tibble")
  require(knitr, include.only = "kable")
  require(ggplot2)
  require(tidyr, include.only = "unnest")
}
```

```{r}
#| label: data_import
tidy_data_path <- params[["tidy_data"]]
d0 <- tidy_data_path |>
  readr::read_rds() |>
  mutate(
    umccrId = glue("{.data$individualId}_{.data$libraryId}_{.data$lane}"),
    umccrId = factor(.data$umccrId),
    projectOwnerName = glue("{.data$projectOwner}_{.data$projectName}")
  ) |>
  select(
    "umccrId",
    "subjectId",
    "libraryId",
    "projectOwnerName",
    "sampleType",
    "phenotype",
    "source",
    "quality",
    "assay",
    "workflow",
    "portalRunId",
    "date",
    "data_tidy"
  )
```

```{r}
#| label: data_setup
ggplot2::theme_set(ggplot2::theme_bw())
options(scipen = 999) # disable scientific notation
options(width = 150)

d_unnest <- d0 |>
  select(
    "umccrId", "libraryId", "subjectId",
    type = "sampleType", "phenotype", "source",
    "quality", "assay", "workflow", "projectOwnerName", "portalRunId", tidy = "data_tidy"
  ) |>
  tidyr::unnest("tidy", names_sep = "_")
# tablename-specific access
# columns: tidy_name, data
# rows: 1 per tidy_table name
d_name <- d_unnest |>
  tidyr::nest(.by = "tidy_name", .key = "data")

## A tibble: 18 × 2
#   tidy_name                     data
#   <glue>                        <list>
# 1 contigmeancov_wgs             <tibble [4 × 4]>
# 2 covmetrics_wgs                <tibble [4 × 4]>
# 3 finehist_wgs                  <tibble [4 × 4]>
# 4 fqc_positionalBaseContent     <tibble [4 × 4]>
# 5 fqc_positionalBaseMeanQuality <tibble [4 × 4]>
# 6 fqc_positionalQuality         <tibble [4 × 4]>
# 7 fqc_readGCContent             <tibble [4 × 4]>
# 8 fqc_readGCContentQuality      <tibble [4 × 4]>
# 9 fqc_readLengths               <tibble [4 × 4]>
# 10 fqc_readMeanQuality           <tibble [4 × 4]>
# 11 fqc_sequencePositions         <tibble [4 × 4]>
# 12 fraglen                       <tibble [3 × 4]>
# 13 hist_wgs                      <tibble [4 × 4]>
# 14 mapmetrics                    <tibble [4 × 4]>
# 15 ploidymetrics                 <tibble [3 × 4]>
# 16 replay                        <tibble [4 × 4]>
# 17 timemetrics                   <tibble [4 × 4]>
# 18 trimmermetrics                <tibble [4 × 4]>

# sample-specific access
# columns: umccrId, data
# rows: 1 per umccrId
d_samp <- d_unnest |>
  tidyr::nest(.by = "umccrId", .key = "data") |>
  arrange(desc(.data$umccrId))
## A tibble: 4 × 2
#  umccrId           data
#  <fct>             <list>
# 1 SBJ05890_L2401624 <tibble [18 × 4]>
# 2 SBJ05889_L2401644 <tibble [16 × 4]>
# 3 SBJ05888_L2401621 <tibble [18 × 4]>
# 4 SBJ05856_L2401572 <tibble [18 × 4]>

dr_unnest <- function(x1, ...) {
  d_name |>
    dplyr::filter(.data$tidy_name == x1) |>
    tidyr::unnest("data") |>
    dplyr::mutate(nrows = purrr::map_int(.data$tidy_data, nrow)) |>
    dplyr::filter(.data$nrows > 0) |>
    dplyr::select(dplyr::everything(), -c("tidy_name", "nrows")) |>
    dplyr::relocate("tidy_data", .after = dplyr::last_col()) |>
    tidyr::unnest("tidy_data")
}
```

```{r}
#| label: funcs
dt_view <- function(x, caption = NULL, scroll_y = 10 + min(nrow(x) * 35, 570), fixedcols = 2, ...) {
  x |>
    DT::datatable(
      filter = list(position = "top", clear = FALSE, plain = TRUE),
      class = "cell-border display compact",
      rownames = FALSE,
      extensions = c("Scroller", "Buttons", "KeyTable", "FixedColumns"),
      options = list(
        scroller = TRUE, scrollY = scroll_y, scrollX = TRUE,
        autoWidth = FALSE, keys = TRUE,
        buttons = c("csv"), dom = "Blfrtip",
        fixedColumns = list(leftColumns = fixedcols)
      ),
      ...
    )
}

blank_lines <- function(n = 10) {
  cat(rep("&nbsp;  ", n), sep = "\n")
}

get_lib_url <- function(lid, text, colour = NULL) {
  url <- glue("https://orcaui.umccr.org/lab?tab=library&search={lid}")
  if (!is.null(colour)) {
    return(glue("<a href={url} style='background-color:{colour}'>{text}</a>"))
  }
  return(glue("<a href={url}>{text}</a>"))
}

type_col <- list(
  wgs = "magenta",
  wts = "purple"
)
```

## Sample Metadata

```{r}
#| label: meta
d0 |>
  arrange(desc(.data$umccrId), libraryId, sampleType) |>
  mutate(
    umccrId = get_lib_url(lid = .data$libraryId, text = .data$umccrId),
    date_analysis_end = lubridate::ymd_hms(.data$date)
  ) |>
  select(umccrId, type = "sampleType", everything(), -data_tidy, -date) |>
  dt_view(escape = FALSE) |>
  DT::formatStyle(
    "type",
    color = DT::styleEqual(
      c("WGS", "WTS"), with(type_col, c(wgs, wts))
    )
  )
```

## QC Metrics

### Mapping

```{r}
#| label: mapmetrics
d_map <- dr_unnest("mapmetrics") |>
  filter(RG == "Total") |>
  arrange(desc(umccrId), type) |>
  mutate(
    umccrId = get_lib_url(lid = .data$libraryId, text = .data$umccrId),
    RG = ifelse(.data$RG == "Total", "Total", "RG")
  ) |>
  select(
    umccrId, subjectId,
    phenotype, type,
    source, quality, assay, workflow, projectOwnerName, portalRunId, RG,
    tot_reads = reads_tot_rg,
    dup_reads = reads_num_dupmarked,
    `dup%` = reads_num_dupmarked_pct,
    `map%` = reads_mapped_pct,
    `unmap%` = reads_unmapped_pct,
    `uniq%` = reads_num_uniq_pct,
    `uniq_map%` = reads_num_uniq_mapped_pct,
    `paired%` = reads_paired_pct,
    `paired_proper%` = reads_paired_proper_pct,
    `singleton%` = reads_singleton_pct,
    `discordant%` = reads_discordant_pct,
    `rrna_filt%` = reads_rrna_filtered_pct,
    `splicejunc%` = reads_splicejunc_pct,
    `mapq_0-10%` = reads_mapq_0_10_pct,
    `mapq_10-20%` = reads_mapq_10_20_pct,
    `mapq_20-30%` = reads_mapq_20_30_pct,
    `mapq_30-40%` = reads_mapq_30_40_pct,
    `mapq_40-Inf%` = reads_mapq_40_inf_pct,
    read_len = read_len,
    insert_len_med = insert_len_median,
    insert_len_mean = insert_len_mean,
    everything(),
    -c("libraryId", "tidy_prefix", "dragen_sample")
  )
num_cols <- purrr::map_lgl(d_map, is.numeric)
num_pct_cols <- grepl("%", names(d_map)) & num_cols
no_numpct_cols <- !grepl("%", names(d_map)) & num_cols
conf <- list(
  bg_size = "95% 90%",
  bg_repeat = "no-repeat",
  bg_pos = "center",
  green_range = c(80, 100),
  pink_range = c(8, 20)
)
d_map |>
  dt_view(fixedcols = 4, scroll_y = 1500, escape = FALSE) |>
  DT::formatCurrency(columns = names(d_map)[num_pct_cols], currency = "", digits = 1) |>
  DT::formatCurrency(columns = names(d_map)[no_numpct_cols], currency = "", digits = 0) |>
  DT::formatStyle(
    columns = c("map%", "uniq%", "uniq_map%", "paired%", "paired_proper%"),
    background = DT::styleColorBar(conf$green_range, color = "lightgreen"),
    backgroundSize = conf$bg_size,
    backgroundRepeat = conf$bg_repeat,
    backgroundPosition = conf$bg_pos
  ) |>
  DT::formatStyle(
    columns = c("dup%", "unmap%", "singleton%", "discordant%"),
    background = DT::styleColorBar(conf$pink_range, color = "pink"),
    backgroundSize = conf$bg_size,
    backgroundRepeat = conf$bg_repeat,
    backgroundPosition = conf$bg_pos
  ) |>
  DT::formatStyle(
    "type",
    color = DT::styleEqual(
      c("WGS", "WTS"), with(type_col, c(wgs, wts))
    )
  )
```

### Coverage + Ploidy

- Ploidy metrics only for **WGS**.

```{r}
#| label: covmetrics
d_pl <- dr_unnest("ploidymetrics") |>
  arrange(desc(umccrId))
d_pl_metrics <- d_pl |>
  select(
    umccrId, subjectId,
    phenotype, type,
    source, quality, assay, workflow, projectOwnerName, portalRunId,
    ploidy = ploidy_est,
    cvg_auto_med_ploidy = cov_autosomal_median,
    cvg_x_med_ploidy = cov_x_median,
    cvg_y_med_ploidy = cov_y_median
  )
# cov_genome_pct_* metrics are in the Hist data, so filter out here
d_cvg <- dr_unnest("covmetrics_wgs") |>
  arrange(desc(umccrId)) |>
  left_join(d_pl_metrics, by = c(
    "umccrId", "subjectId", "phenotype", "type", "source",
    "quality", "assay",
    "workflow", "projectOwnerName", "portalRunId"
  )) |>
  mutate(umccrId = get_lib_url(lid = .data$libraryId, text = .data$umccrId)) |>
  select(
    "umccrId", "subjectId", "phenotype", "type", "source",
    "quality", "assay",
    "workflow", "projectOwnerName", "portalRunId",
    ploidy,
    cvg_auto_avg = cov_avg_auto_over_genome,
    cvg_auto_med = cov_median_auto_over_genome,
    cvg_x_avg = cov_avg_x_over_genome,
    cvg_y_avg = cov_avg_y_over_genome,
    cvg_uniq = cov_alignment_avg_over_genome,
    cvg_mito_avg = cov_avg_mt_over_genome,
    cvg_auto_med_ploidy,
    cvg_x_med_ploidy,
    cvg_y_med_ploidy,
    reads_aligned_tot,
    bases_aligned_tot,
    cvg_gt02 = cov_uniformity_pct_gt02mean_genome,
    cvg_gt04 = cov_uniformity_pct_gt04mean_genome,
    everything(),
    -contains("cov_pct_"),
    -c("libraryId", "tidy_prefix")
  )
num_cols <- names(d_cvg)[purrr::map_lgl(d_cvg, is.numeric)]
d_cvg |>
  dt_view(fixedcols = 4, scroll_y = 1500, escape = FALSE) |>
  DT::formatCurrency(columns = num_cols, currency = "", digits = 1) |>
  DT::formatStyle(
    "ploidy",
    backgroundColor = DT::styleEqual(
      c("XY", "XX"), c("lightblue", "lightpink")
    )
  ) |>
  DT::formatStyle(
    "type",
    color = DT::styleEqual(
      c("WGS", "WTS"), with(type_col, c(wgs, wts))
    )
  )
```

### Trimmer

```{r}
#| label: trimmermetrics
d_tr <- dr_unnest("trimmermetrics") |>
  arrange(desc(umccrId), type) |>
  mutate(
    umccrId = get_lib_url(lid = .data$libraryId, text = .data$umccrId)
  ) |>
  select(
    umccrId, subjectId,
    phenotype, type,
    source, quality, assay, workflow, projectOwnerName, portalRunId,
    reads_tot = reads_tot_input,
    read_len_avg = read_len_avg,
    `polygkmers3r1_remain%` = polygkmers3r1_remaining_pct,
    `polygkmers3r2_remain%` = polygkmers3r2_remaining_pct,
    `polyg_soft_trimmed_reads_unfilt_3r1%` = polyg_soft_trimmed_reads_unfilt_3r1_pct,
    `polyg_soft_trimmed_reads_unfilt_3r2%` = polyg_soft_trimmed_reads_unfilt_3r2_pct,
    `polyg_soft_trimmed_bases_unfilt_3r1%` = polyg_soft_trimmed_bases_unfilt_3r1_pct,
    `polyg_soft_trimmed_bases_unfilt_3r2%` = polyg_soft_trimmed_bases_unfilt_3r2_pct,
    polygkmers3r1_remaining = polygkmers3r1_remaining,
    polygkmers3r2_remaining = polygkmers3r2_remaining,
    polyg_soft_trimmed_reads_unfilt_3r1 = polyg_soft_trimmed_reads_unfilt_3r1,
    polyg_soft_trimmed_reads_unfilt_3r2 = polyg_soft_trimmed_reads_unfilt_3r2,
    polyg_soft_trimmed_bases_unfilt_3r1 = polyg_soft_trimmed_bases_unfilt_3r1,
    polyg_soft_trimmed_bases_unfilt_3r2 = polyg_soft_trimmed_bases_unfilt_3r2,
    bases_tot = bases_tot,
    bases_r1 = bases_r1,
    bases_r2 = bases_r2,
    reads_trimmed_tot = reads_trimmed_tot,
    `reads_trimmed_tot%` = reads_trimmed_tot_pct,
    bases_trimmed_tot = bases_trimmed_tot,
    `bases_trimmed_tot%` = bases_trimmed_tot_pct,
    reads_tot_filt = reads_tot_filt,
    `reads_tot_filt%` = reads_tot_filt_pct,
    everything(),
    -c("libraryId", "tidy_prefix")
  )
d_tr |>
  dt_view(fixedcols = 4, escape = FALSE) |>
  DT::formatStyle(
    "type",
    color = DT::styleEqual(
      c("WGS", "WTS"), with(type_col, c(wgs, wts))
    )
  )
```

### RunTime Metrics

```{r}
#| label: timemetrics
#| fig-height: 13
d_time <- dr_unnest("timemetrics")
d_time_plot <- d_time |>
  select(umccrId, type, phenotype, TotalRuntime:dplyr::last_col()) |>
  tidyr::pivot_longer(TotalRuntime:dplyr::last_col(), names_to = "STEP", values_to = "RUNTIME_SEC") |>
  filter(!is.na(.data$RUNTIME_SEC)) |>
  ggplot(aes(x = RUNTIME_SEC, y = STEP, colour = umccrId)) +
  geom_point() +
  facet_wrap(~type, ncol = 1, scales = "free")

plotly::ggplotly(d_time_plot)
```


## FastQC {.tabset .tabset-pills}

- Plot names used in MultiQC report are shown in **parentheses** for easier comparison.
- For interactive plots, focus on a specific sample by **double-clicking** its label
in the legend.

### Read Mean Quality ('Per-Sequence Quality Scores')

```{r}
#| label: fqc_readMeanQuality
#| fig-height: 10
f1 <- dr_unnest("fqc_readMeanQuality") |>
  group_by(umccrId, mate) |>
  mutate(
    tot = sum(.data$value),
    prop = round(.data$value / .data$tot, 3),
    prop = 100 * prop
  ) |>
  ungroup()

fqc_colours1 <- tibble::tibble(
  start = c(0, 20, 28),
  end = c(20, 28, max(f1$q) + 2),
  col = c("#e6c3c3", "#e6dcc3", "#c3e6c3")
)
f1_plot <- ggplot() +
  geom_rect(
    data = fqc_colours1,
    mapping = aes(xmin = start, xmax = end, ymin = 0, ymax = max(f1$prop) + 2),
    fill = rep(fqc_colours1$col, length(unique(f1$type))),
    alpha = 0.7
  ) +
  geom_line(data = f1, aes(x = q, y = prop, colour = umccrId, linetype = mate), linewidth = 1, show.legend = FALSE) +
  scale_y_continuous(labels = scales::label_comma()) +
  theme(panel.grid.major = element_blank()) +
  facet_wrap(~type, ncol = 1) +
  labs(
    title = "Read Mean Quality",
    subtitle = glue("Percentage of reads with average quality scores. Shows if\na subset of reads has poor quality.")
  )

# plotly::ggplotly(f1_plot)
f1_plot
```

### GC Content ('Per-Sequence GC Content')

```{r}
#| label: fqc_readGCContent
#| fig-height: 10
gc_data <- dr_unnest("fqc_readGCContent") |>
  group_by(umccrId, mate) |>
  mutate(
    tot = sum(.data$value),
    prop = round(.data$value / .data$tot, 3),
    prop = 100 * prop
  ) |>
  ungroup()

gc_data_plot <- gc_data |>
  ggplot(aes(x = pct, y = prop, colour = umccrId)) +
  geom_line(aes(linetype = mate), alpha = 0.4, linewidth = 1) +
  facet_wrap(~type, ncol = 1) +
  labs(
    x = "GC content",
    y = "Percentage of Reads",
    title = "Read GC Content",
    subtitle = glue("Total number of reads with each GC content\npercentile between 0% and 100%")
  )
plotly::ggplotly(gc_data_plot)
# gc_data_plot
```

### GC Content Quality ('GC Content Mean Quality Scores')

```{r}
#| label: fqc_readGCContentQuality
f1 <- dr_unnest("fqc_readGCContentQuality") |>
  filter(!is.na(.data$value))
fqc_colours2 <- tibble::tibble(
  start = c(0, 20, 28),
  end = c(20, 28, max(f1$value) + 2),
  col = c("#e6c3c3", "#e6dcc3", "#c3e6c3")
)
f1_plot <- ggplot() +
  geom_rect(
    data = fqc_colours2,
    mapping = aes(ymin = start, ymax = end, xmin = 0, xmax = 100),
    fill = rep(fqc_colours2$col, length(unique(f1$type))),
    alpha = 0.7
  ) +
  geom_line(data = f1, aes(x = pct, y = value, colour = umccrId, linetype = mate), linewidth = 1) +
  facet_wrap(~type, ncol = 1) +
  labs(
    title = "GC Content Quality",
    subtitle = glue("Average Phred-scale read mean quality for reads with\neach GC content percentile between 0% and 100%.")
  )
plotly::ggplotly(f1_plot)
# f1_plot
```

### Positional Base Content ('Per-Position Sequence Content')

- TODO: create heatmap instead

```{r}
#| label: fqc_pbc
#| fig-height: 42
#| eval: false
f1 <- dr_unnest("fqc_positionalBaseContent")
f1 |>
  filter(base != "N") |>
  mutate(prop = prop * 100) |>
  ggplot(aes(x = pos, y = prop, colour = base, group = base)) +
  geom_line() +
  facet_grid(forcats::fct_rev(umccrId) ~ mate) +
  labs(
    x = "Position in Read (bp)",
    y = "Proportion of Bases",
    title = "Positional Base Content",
    subtitle = glue(
      "Proportion of bases of each specific nucleotide",
      "\nat given locations in the read."
    )
  )
```

### Positional Base Mean Quality ('Per-Position Mean Quality Scores')


```{r}
#| label: fqc_bmq
#| fig-height: 80
#| eval: false
f1 <- dr_unnest("fqc_positionalBaseMeanQuality")
ggplot() +
  geom_rect(
    data = fqc_colours2,
    mapping = aes(ymin = start, ymax = end, xmin = -Inf, xmax = Inf),
    fill = rep(fqc_colours2$col, length(unique(f1$umccrId)) * length(unique(f1$mate))),
    alpha = 0.7
  ) +
  geom_line(data = f1, aes(x = pos, y = value, colour = base)) +
  facet_grid(forcats::fct_rev(umccrId) ~ mate) +
  labs(
    x = "Position in Read (bp)",
    y = "Quality Score",
    title = "Positional Base Mean Quality",
    subtitle = glue(
      "Average Phred-scale quality value of bases with a specific",
      "\nnucleotide and at a given location in the read."
    )
  )
```

### Positional Quality ('Per-Position Quality Score Ranges')

```{r}
#| label: fqc_pq
#| fig-width: 13
#| eval: false

# TODO: use boxplot instead of point
f1 <- dr_unnest("fqc_positionalQuality")
quants <- c(25, 50, 75)
f1 |>
  mutate(pos = as.integer(.data$pos)) |>
  filter(pct %in% quants) |>
  ggplot(aes(x = pos, y = value, colour = pct)) +
  geom_point() +
  facet_wrap(~ forcats::fct_rev(umccrId)) +
  labs(
    title = "Positional Quality",
    subtitle = glue("Phred-scale quality value for bases at a given location and a\ngiven quantile of the distribution ({paste(quants, collapse = ', ')})")
  )
```

### Read Lengths ('Sequence Length Distribution')

```{r}
#| label: fqc_readLengths
#| fig-height: 10
read_len <- dr_unnest("fqc_readLengths")
read_len_plot <- read_len |>
  group_by(umccrId, mate) |>
  mutate(
    tot = sum(.data$value),
    prop = round(.data$value / .data$tot, 3),
    prop = 100 * prop
  ) |>
  ungroup() |>
  select(umccrId, type, mate, bp, value, tot, prop) |>
  ggplot(aes(x = bp, y = prop, colour = umccrId, shape = mate)) +
  geom_jitter(width = 0.1, height = 0.1) +
  # geom_line(aes(linetype = mate), linewidth = 1) +
  theme(
    panel.grid.major = element_blank()
  ) +
  facet_wrap(~type, ncol = 1) +
  labs(
    title = "Read Lengths",
    subtitle = glue("Read percentage with each observed length.")
  )
plotly::ggplotly(read_len_plot)
# read_len_plot
```

### Sequence Positions ('Adapter Content')

```{r}
#| label: fqc_sequencePositions
#| fig-height: 80
#| eval: false
f1 <- dr_unnest("fqc_sequencePositions")
f1 |>
  ggplot(aes(x = bp, y = value, colour = seq)) +
  geom_line() +
  facet_grid(forcats::fct_rev(umccrId) ~ mate, scales = "free_y") +
  labs(title = glue(
    "Number of times an adapter or other kmer sequence is found,\n",
    "starting at a given position in the input reads."
  ))
```

---

## Coverage {.tabset .tabset-pills}

```{r}
#| label: contig_cvg
#| fig-height: 5
#| eval: false
#| results: asis
# TODO: FIXME
d1 <- dr_unnest("contigmeancov_wgs") |>
  arrange(desc("umccrId"))
for (type1 in sort(unique(d1$type), decreasing = FALSE)) {
  cat(glue("\n\n### {type1} {{.tabset .tabset-pills}}"), "\n\n")
  d1_type <- d1 |>
    filter(type == type1)
  for (s in sort(unique(d1_type$umccrId), decreasing = TRUE)) {
    p1 <- d1_type |>
      filter(umccrId == s) |>
      dracarys::WgsContigMeanCovFile$public_methods$plot() +
      labs(subtitle = s)
    cat(glue("\n#### {s}"), "\n")
    blank_lines(2)
    print(p1)
    blank_lines(2)
    cat("\n\n\n")
  }
  cat("\n\n\n")
}
```

-----

## Fragment Length

- Only for WGS.

```{r}
#| label: fraglenhist_plot
#| fig-height: 8

fl1 <- dr_unnest("fraglen")
min_count <- 10
flp <- fl1 |>
  filter(.data$count >= min_count) |>
  ggplot(aes(x = .data$fragmentLength, y = .data$count)) +
  geom_line(aes(colour = umccrId)) +
  labs(title = "Fragment Length Distribution") +
  xlab("Fragment Length (bp)") +
  ylab(glue("Read Count (min: {min_count})")) +
  theme(
    legend.position.inside = c(0.9, 0.9),
    legend.justification = c(1, 1),
    panel.grid.minor = ggplot2::element_blank(),
    plot.title = ggplot2::element_text(colour = "#2c3e50", size = 14, face = "bold")
  )
plotly::ggplotly(flp)
# flp
```

---

- Only for WGS.

```{r pe, eval=T, fig.height=5}
#| label: pe
#| eval: FALSE
#| fig-height: 5

chrom_levels <- c(1:22, "x", "y")
d_pl_plot_data <- d_pl |>
  select(
    umccrId, phenotype, type,
    contains("div_auto_median")
  ) |>
  tidyr::pivot_longer(-c("umccrId", "phenotype", "type")) |>
  tidyr::separate_wider_delim("name", delim = "_", names = c("cov", "chrom", "rest"), too_many = "merge") |>
  mutate(chrom = factor(chrom, levels = chrom_levels)) |>
  select(umccrId, phenotype, type, chrom, value)

d_pl_plot <- d_pl_plot_data |>
  ggplot(aes(x = chrom, y = value)) +
  geom_line(aes(colour = umccrId, group = umccrId), na.rm = TRUE) +
  geom_point(aes(colour = umccrId), na.rm = TRUE) +
  labs(title = "Chromosome Median / Autosomal Median")
plotly::ggplotly(d_pl_plot)
# d_pl_plot
```


---

## Hist

```{r}
#| label: cvgm
#| eval: false
#| fig-height: 8
#| fig-width: 12
d_hist <- dr_unnest("hist_wgs")
d_hist1 <- d_hist |>
  ggplot(aes(x = start, y = pct, colour = umccrId)) +
  geom_point() +
  geom_linerange(aes(xmin = start, xmax = end)) +
  scale_y_continuous(n.breaks = 10) +
  scale_x_continuous(n.breaks = 10) +
  facet_wrap(~type, nrow = 1) +
  labs(
    x = "Coverage range", y = "PCT",
    title = "PCT of bases with coverage in the given range",
    subtitle = "e.g. X PCT of bases have coverage between 100 and 500."
  )
d_hist2 <- d_hist |>
  ggplot(aes(x = start, y = cumsum, colour = umccrId)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(n.breaks = 10) +
  facet_wrap(~type, nrow = 1) +
  labs(
    x = "Coverage", y = "PCT Cumsum",
    title = "Percentage of sites in genome with given coverage"
  )
# plotly::subplot(d_hist1, d_hist2, shareY = TRUE, titleY = TRUE, titleX = TRUE, nrows = 2)
```

## FineHist

```{r}
#| label: finehist
#| eval: false
#| fig-height: 10
#| fig-width: 12
d_fhist <- dr_unnest("finehist_wgs")
d_fhist |>
  dracarys::WgsFineHistFile$public_methods$plot(c(0, 150)) +
  facet_wrap(~ forcats::fct_rev(umccrId), scales = "free_y")
```

---
